---
title: "Deliberative Ignorance"
author: "Eva Crone"
date: "2025-10-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
```

# 1. Introduction

This document explores the DL2_preprocessed dataset to gain an initial understanding of the data and investigate patterns in deliberate ignorance (DI) across scenarios, motives, and age groups.

# 2. Data Preparation
## 2.1 Load Data
```{r}
dat <- read_csv("dl2_data_processed.csv")
```

## 2.2 Initial Cleaning
Exclude participants with missing DI Ratings.
```{r}
dat_short <- dat %>%
filter(!is.na(want_to_know_bin))
```

## 2.3. Include categories in dat_short
```{r}
dat_short <- dat_short %>%
  mutate(
    Scenario_Category = case_when(
      Scenario_Content %in% c(
        "hiv_test", "gen_risk_alzheimer", "gen_risk_huntington", 
        "life_expect_huntington", "spread_metast_cancer", "risk_anaesthesia",
        "nutritional_deficiencies", "calories_food", "sex_of_child"
      ) ~ "Health",
      
      Scenario_Content %in% c(
        "partner_cheating", "partner_prev_relationship",
        "converts",
        "friend_polit_opin", "friend_sex_life"
      ) ~ "Social",
      
      Scenario_Content %in% c(
        "colleague_salary", "bank_balance",
        "altern_recent_purchase", "friend_finance_problems"
      ) ~ "Finances",
      
      Scenario_Content %in% c(
        "success_chance_new_comp", "gender_job_applicant", "teaching_eval"
      ) ~ "Career",
      
      Scenario_Content %in% c(
        "hunger_crisis", "war", "holocaust_movie",
        "child_abuse", "homeless_ppl"
      ) ~ "Society",
      
      Scenario_Content %in% c(
        "Kardashian_home", "ending_GoT"
      ) ~ "Entertainment",
      
      Scenario_Content %in% c(
        "global_warming", "factory_farming"
      ) ~ "Environment",
    )
  )

```


## 2.4. Create Wide Format 
```{r}
motive_vars <- c("expectation","valence","arousal", "sadness","anger","fear", 
                  "jealousy","envy","disgust","regret","feelinggood","relief",
                  "happiness","complexity","relevance","reliability","relationships",
                  "conflict","politeness","notlying","unsureresp.","exclusion",
                  "self.percept.","actiongeneral","actionnow","disadvantage",
                  "goalinterfer","socialobligat","legalobligat","surprise","fairness")


data_wide <- dat_short %>%
  select(ResponseId, Age, Age_group, Scenario_Code, all_of(motive_vars)) %>%
  # long -> wide
  pivot_wider(
    id_cols = c(ResponseId, Age, Age_group), 
    names_from = Scenario_Code,
    values_from = all_of(motive_vars),
    names_sep = "_"
  )

```


# 3. Participants Overview 

## 3.1. Participants in total
```{r}
nrow(data_wide)
```

## 3.2. Participants count young and old

```{r}
data_wide %>%
  dplyr::group_by(Age_group) %>%
  dplyr::summarise(count = n())
```

## 3.3. Age range 
```{r}
range(data_wide$Age, na.rm = TRUE)
```
### Plot Age range
```{r}
hist(data_wide$Age,
     main = "Age range",
     xlab = "Age",
     ylab = "Frequency",
     col = "skyblue",
     border = "white")
```


## 3.4. Ratings per participant

### Number of DI ratings per participant
```{r}
ratings_count_person <- dat_short %>%
  dplyr::filter(want_to_know_bin %in% c(0, 1)) %>%  
  dplyr::group_by(ResponseId) %>%
  dplyr::summarise(
    n_valid_ratings = n(),   
    .groups = "drop"
  )
```

### Number of Motive Ratings per Participant
```{r}
rating_cols <- setdiff(names(data_wide), c("ResponseId", "Age_group"))

motives_count_person <- data_wide %>%
  mutate(
    n_ratings_filled = rowSums(
      select(., all_of(rating_cols)) != -99 & !is.na(select(., all_of(rating_cols)))
    )
  ) %>%
  select(ResponseId, n_ratings_filled)

### Combine Both Counts
ratings_combined <- merge(
  ratings_count_person,
  motives_count_person,
  by = "ResponseId"
)

ratings_combined <- ratings_combined %>%
  mutate(avg_ratings_per_scenario = n_ratings_filled / n_valid_ratings)
```

### Overall mean rating per scenario
```{r}
mean(ratings_combined$avg_ratings_per_scenario, na.rm = TRUE)
```


## 3.5. Number of ratings by scenario and age group
```{r}
dat_short <- dat_short %>%
  mutate(Age_group = ifelse(is.na(Age_group), "unknown", Age_group))

scenario_participants <- dat_short %>%
  dplyr::group_by(Scenario_Content, Age_group) %>%
  dplyr::summarise(n_ratings = n(), .groups = "drop")

scenario_participants
```

# 4. Deliberative Ignorance 
Responses Want to Know / Not Want to Know (DI) 

## 4.1 Overall proportion of DI across all scenarios
```{r}
overall_DI <- dat_short %>%
  dplyr::filter(!is.na(want_to_know_bin)) %>%
  dplyr::summarise(
    total_responses = n(),
    total_DI = sum(want_to_know_bin == 1),
    proportion_DI = mean(want_to_know_bin == 1)
  )

overall_DI 
```

## 4.2. DI Ratings by scenario 
### How many participants in general answered the scenario
```{r}
scenario_counts <- dat_short %>%
  dplyr::group_by(Scenario_Content) %>%
  dplyr::summarise(n_ratings = n(), .groups = "drop")

scenario_counts
```


### Count of want to know = 1 (DI) per scenario
```{r}
di_scenario <- dat_short %>% 
  dplyr::group_by(Scenario_Content, want_to_know_bin) %>% 
  dplyr::summarise(n_ratings = n(),.groups = "drop")

```

### Plot DI responses by scenario
```{r pressure, echo=FALSE}
ggplot(di_scenario, aes(x = Scenario_Content, y = n_ratings, fill = factor(want_to_know_bin))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("0" = "skyblue", "1" = "#E69F00"),
                    name = "Want to know?",
                    labels = c("0 = want to know", "1 = do NOT want to know (DI)")) +
  labs(
    title = "Want-to-Know Responses by Scenario",
    x = "Scenario",
    y = "Number of Ratings"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold", hjust = 0.5)
  )

```


## 4.3. Proportion DI vs. no DI by scenario
```{r}
di_scenario_full <- dat_short %>% 
  dplyr::group_by(Scenario_Content) %>%
  dplyr::summarise(
    n_0  = sum(want_to_know_bin == 0, na.rm = TRUE),
    n_di  = sum(want_to_know_bin == 1, na.rm = TRUE),
    total = n_0 + n_di,
    prop_1 = n_di / total,
    .groups = "drop"
  )%>%
  dplyr::arrange(desc(prop_1))

di_scenario_full <- di_scenario_full %>%
  mutate(Scenario_Content = factor(Scenario_Content, levels = Scenario_Content[order(prop_1)]))
```

### Plot proportion DI vs. no DI by scenario
```{r}
ggplot(di_scenario_full, aes(x = Scenario_Content, y = prop_1)) +
  geom_col(fill = "#E69F00") +
  coord_flip() +
  labs(
    title = "Proportion of DI per Scenario",
    x = "Scenario",
    y = "Proportion of DI to total"
  ) +
  theme_minimal(base_size = 14)
```

## 4.4. Count DI vs. no DI by category 
```{r}
di_category <- dat_short %>% 
  dplyr::group_by(Scenario_Category, want_to_know_bin) %>% 
  dplyr::summarise(n_ratings = n(),.groups = "drop")
```

### Plot DI count by category
```{r}
ggplot(di_category, aes(x = Scenario_Category, y = n_ratings, fill = factor(want_to_know_bin))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("0" = "skyblue", "1" = "#E69F00"),
                    name = "Want to know?",
                    labels = c("0 = want to know", "1 = do NOT want to know (DI)")) +
  labs(
    title = "Want-to-Know Responses by Category",
    x = "Category",
    y = "Number of Ratings"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold", hjust = 0.5)
  )

```


## 4.5.Proportion DI by category
```{r}
di_category_full <- dat_short %>% 
  dplyr::group_by(Scenario_Category) %>%
  dplyr::summarise(
    n_0  = sum(want_to_know_bin == 0, na.rm = TRUE),
    n_di  = sum(want_to_know_bin == 1, na.rm = TRUE),
    total = n_0 + n_di,
    prop_1 = n_di / total,
    .groups = "drop"
  )%>%
  dplyr::arrange(desc(prop_1))

di_category_full <- di_category_full %>%
  mutate(Scenario_Category = factor(Scenario_Category, levels = Scenario_Category[order(prop_1)]))
```

### Plot proportion DI vs. no DI by category
```{r}
ggplot(di_category_full, aes(x = Scenario_Category, y = prop_1)) +
  geom_col(fill = "#E69F00") +
  coord_flip() +
  labs(
    title = "Proportion of DI per Category",
    x = "Category",
    y = "Proportion of DI to total"
  ) +
  theme_minimal(base_size = 14)
```

# 5. Deliberative Ignorance — young vs. old 
Decisions Want to Know / Not to Know (DI)

## 5.1. Difference overall proportion -- young vs. old 
```{r}
overall_DI_age <- dat_short %>%
  dplyr::filter(!is.na(want_to_know_bin)) %>%
  dplyr::group_by(Age_group) %>%        
  dplyr::summarise(
    total_responses = n(),
    total_DI = sum(want_to_know_bin == 1),
    proportion_DI = mean(want_to_know_bin == 1),
    .groups = "drop"
  )
overall_DI_age
```

## 5.2. Differences DI by scenario --  young vs. old 
positive: more young participants chose “not want to know” 
```{r}
scenario_age_prop <- dat_short %>%
  dplyr::group_by(Scenario_Content, Age_group) %>%
  dplyr::summarise(
    total_participants = n(),
    n_want_to_know = sum(want_to_know_bin == 1, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::mutate(prop_want_to_know = n_want_to_know / total_participants) %>%
  dplyr::select(-n_want_to_know, -total_participants) %>%
  pivot_wider(
    names_from = Age_group,
    values_from = prop_want_to_know,
    values_fill = 0
  ) %>%
  mutate(diff_prop_young_old = younger - older) %>%
  arrange(desc(diff_prop_young_old)) 

scenario_age_prop 
```


### Plot Differences DI by scenario and age group

```{r}
scenario_age_long <- scenario_age_prop %>%
  pivot_longer(cols = c(younger, older),
               names_to = "Age_group",
               values_to = "prop_want_to_know") %>%
  group_by(Scenario_Content) %>%
  mutate(mean_prop = mean(prop_want_to_know, na.rm = TRUE)) %>%
  ungroup()

ggplot(scenario_age_long,
       aes(x = reorder(Scenario_Content, mean_prop),
           y = prop_want_to_know,
           fill = Age_group)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("younger" = "#4CAF50", "older" = "#2196F3"),
                    labels = c("older", "younger")) +
  labs(
    title = "DI ratings by scenario and age group",
    x = "Scenario",
    y = "Proportion DI",
    fill = "Age group"
  ) +
  coord_flip() +
  theme_minimal(base_size = 13)
```


## 5.3. Differences in DI by category -- young vs. old
positive: more young participants chose “not want to know” 
```{r}
category_age_prop <- dat_short %>%
  dplyr::group_by(Scenario_Category, Age_group) %>%
  dplyr::summarise(
    total_participants = n(),
    n_want_to_know = sum(want_to_know_bin == 1, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::mutate(prop_want_to_know = n_want_to_know / total_participants) %>%
  dplyr::select(-n_want_to_know, -total_participants) %>%
  pivot_wider(
    names_from = Age_group,
    values_from = prop_want_to_know,
    values_fill = 0
  ) %>%
  mutate(diff_prop_young_old = younger - older) %>%
  arrange(desc(diff_prop_young_old)) 

category_age_prop
```

### Plot Differences DI by category and age group

```{r}
scenario_age_long_cat <- category_age_prop %>%
  pivot_longer(cols = c(younger, older),
               names_to = "Age_group",
               values_to = "prop_want_to_know") %>%
  group_by(Scenario_Category) %>%
  mutate(mean_prop = mean(prop_want_to_know, na.rm = TRUE)) %>%
  ungroup()

ggplot(scenario_age_long_cat,
       aes(x = reorder(Scenario_Category, mean_prop),
           y = prop_want_to_know,
           fill = Age_group)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("older" = "#4CAF50", "younger" = "#2196F3"),
                    labels = c("older", "younger")) +
  labs(
    title = "DI ratings by category and age group",
    x = "Category",
    y = "Response Not want to know (DI)",
    fill = "Age group"
  ) +
  coord_flip() +
  theme_minimal(base_size = 13)
```




# 6.Ratings of Motives

## 6.1. Overall highly rated motives
```{r}
overall_motives <- dat_short %>%
  summarise(across(all_of(motive_vars), ~ mean(.x[.x != -99], na.rm = TRUE))) %>%
  pivot_longer(cols = everything(), names_to = "motive", values_to = "mean_rating") %>%
  arrange(desc(mean_rating))

overall_motives
```

## 6.2. Mean Ratings of motives by scenario
```{r}
ratings_per_motive_mean <- dat_short %>%
  dplyr::group_by(Scenario_Content) %>%
  dplyr::summarise(
    across(
      all_of(motive_vars),
      ~ mean(.x[.x != -99], na.rm = TRUE),
      .names = "mean_{.col}"
    ),
    .groups = "drop"
  )
```


### Plot: Heatmap mean ratings of motives by scenario
```{r}
ratings_long <- ratings_per_motive_mean %>%
  pivot_longer(
    cols = starts_with("mean_"),
    names_to = "Motive",
    values_to = "MeanRating"
  )


ggplot(ratings_long, aes(x = Motive, y = Scenario_Content, fill = MeanRating)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "steelblue") +
  labs(
    title = "Mean Ratings of Motives per Scenario",
    x = "Motive",
    y = "Scenario",
    fill = "Mean Rating"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

## 6.3. Top 3 highest average rated motives by scenario
```{r}
top3_motives <- dat_short %>%
  dplyr::group_by(Scenario_Content) %>%
  dplyr::summarise(across(all_of(motive_vars),
                   ~ mean(.x[.x != -99], na.rm = TRUE),  
                   .names = "avg_{.col}"),
            .groups = "drop") %>%
  pivot_longer(
    cols = starts_with("avg_"),
    names_to = "motive",
    values_to = "avg_rating"
  ) %>%
  mutate(motive = sub("^avg_", "", motive)) %>%
  dplyr::group_by(Scenario_Content) %>%
  slice_max(avg_rating, n = 3, with_ties = FALSE) %>% 
  arrange(Scenario_Content, desc(avg_rating)) %>%
  ungroup()
```


### Table: Top 3 average motives per scenario 
```{r}
top3_motives_table <- dat_short %>%
  dplyr::group_by(Scenario_Content) %>%
  dplyr::summarise(
    across(
      all_of(motive_vars),
      ~ mean(.x[.x != -99], na.rm = TRUE),  
      .names = "avg_{.col}"
    ),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = starts_with("avg_"),
    names_to = "motive",
    values_to = "avg_rating"
  ) %>%
  mutate(motive = sub("^avg_", "", motive)) %>%
  dplyr::group_by(Scenario_Content) %>%
  slice_max(avg_rating, n = 3, with_ties = FALSE) %>%
  arrange(Scenario_Content, desc(avg_rating)) %>%
  mutate(rank = row_number()) %>%
  pivot_wider(
    names_from = rank,
    values_from = c(motive, avg_rating),
    names_glue = "Top{rank}_{.value}"
  ) %>%
  arrange(Scenario_Content)

knitr::kable(top3_motives_table, digits = 2, caption = "Top 3 Motives (Mean Ratings) per Scenario")
```

### Plot Top 3 Motives by Scenario

```{r}
ggplot(top3_motives, aes(x = motive, y = Scenario_Content, fill = avg_rating)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "steelblue") +
  labs(
    title = "Top 3 Motives by Scenario",
    x = "Motive",
    y = "Scenario",
    fill = "Mean Rating"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 6.4. Mean Ratings of motives by category 

```{r}
ratings_per_motive_mean_cat <- dat_short %>%
  dplyr::group_by(Scenario_Category) %>%
  dplyr::summarise(
    across(
      all_of(motive_vars),
      ~ mean(.x[.x != -99], na.rm = TRUE),
      .names = "mean_{.col}"
    ),
    .groups = "drop"
  )

ratings_per_motive_mean_cat
```

### Plot: Heatmap mean ratings of motives by category
```{r}
ratings_long_cat <- ratings_per_motive_mean_cat %>%
  pivot_longer(
    cols = starts_with("mean_"),
    names_to = "Motive",
    values_to = "MeanRating"
  )


ggplot(ratings_long_cat, aes(x = Motive, y = Scenario_Category, fill = MeanRating)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "steelblue") +
  labs(
    title = "Mean Ratings of Motives by Category",
    x = "Motive",
    y = "Category",
    fill = "Mean Rating"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```



## 6.5. Top 3 highest average rated motives by category
```{r}
top3_motives_cat <- dat_short %>%
  dplyr::group_by(Scenario_Category) %>%
  dplyr::summarise(across(all_of(motive_vars),
                          ~ mean(.x[.x != -99], na.rm = TRUE),
                          .names = "avg_{.col}"),
                   .groups = "drop") %>%
  pivot_longer(
    cols = starts_with("avg_"),
    names_to = "motive",
    values_to = "avg_rating"
  ) %>%
  mutate(motive = sub("^avg_", "", motive)) %>%
  dplyr::group_by(Scenario_Category) %>%
  slice_max(avg_rating, n = 3, with_ties = FALSE) %>% 
  arrange(Scenario_Category, desc(avg_rating)) %>%
  ungroup()
```

### Table: Top3 motives by category
```{r}
top3_motives_cat_table <- dat_short %>%
  dplyr::group_by(Scenario_Category) %>%
  dplyr::summarise(
    across(
      all_of(motive_vars),
      ~ mean(.x[.x != -99], na.rm = TRUE),  
      .names = "avg_{.col}"
    ),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = starts_with("avg_"),
    names_to = "motive",
    values_to = "avg_rating"
  ) %>%
  mutate(motive = sub("^avg_", "", motive)) %>%
  dplyr::group_by(Scenario_Category) %>%
  slice_max(avg_rating, n = 3, with_ties = FALSE) %>%
  arrange(Scenario_Category, desc(avg_rating)) %>%
  mutate(rank = row_number()) %>%
  pivot_wider(
    names_from = rank,
    values_from = c(motive, avg_rating),
    names_glue = "Top{rank}_{.value}"
  ) %>%
  arrange(Scenario_Category)


knitr::kable(top3_motives_cat_table, digits = 2, caption = "Top 3 Motives (Mean Ratings) per Category")
```


### Plot Top 3 motives by category
```{r}
ggplot(top3_motives_cat, aes(x = motive, y = Scenario_Category, fill = avg_rating)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "steelblue") +
  labs(
    title = "Top 3 Motives by Category",
    x = "Motive",
    y = "Category",
    fill = "Mean Rating"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```



# 7. Ratings of Motives -- young vs. old 

## 7.1. Overall highly rated motives -- young vs. old 
```{r}
overall_motives_by_age <- dat_short %>%
  dplyr::filter(Age_group %in% c("younger", "older")) %>%   
  dplyr::group_by(Age_group) %>%
  dplyr::summarise(
    across(all_of(motive_vars),
           ~ mean(.x[.x != -99], na.rm = TRUE),
           .names = "avg_{.col}")
  ) %>%
  tidyr::pivot_longer(
    cols = starts_with("avg_"),
    names_to = "motive",
    values_to = "mean_rating"
  ) %>%
  dplyr::mutate(motive = sub("^avg_", "", motive)) %>%
  dplyr::arrange(Age_group, desc(mean_rating))

overall_motives_by_age
```

### Plot Overall motive ratings by age group
```{r}
ggplot(overall_motives_by_age, aes(x = reorder(motive, mean_rating), 
                                   y = mean_rating, 
                                   fill = Age_group)) +
  geom_col(position = position_dodge(width = 0.8)) +
  coord_flip() +
  scale_fill_manual(values = c("older" = "#4CAF50", "younger" = "#2196F3"),
                    labels = c("older", "younger")) +
  labs(
    title = "Overall Motive Ratings by Age Group",
    x = "Motive",
    y = "Average Rating",
    fill = "Age Group"
  ) +
  theme_minimal(base_size = 12)
```



## 7.2. Top 3 motives by scenario - young vs. old  
```{r}
top3_motives_by_age <- dat_short %>%
  dplyr::filter(Age_group %in% c("younger", "older")) %>%  
  dplyr::group_by(Scenario_Content, Age_group) %>%
  dplyr::summarise(
    across(all_of(motive_vars),
           ~ mean(.x[.x != -99], na.rm = TRUE),
           .names = "avg_{.col}"),
    .groups = "drop"
  ) %>%
  tidyr::pivot_longer(
    cols = starts_with("avg_"),
    names_to = "motive",
    values_to = "avg_rating"
  ) %>%
  dplyr::mutate(motive = sub("^avg_", "", motive)) %>%
  dplyr::group_by(Scenario_Content, Age_group) %>%
  dplyr::slice_max(avg_rating, n = 3, with_ties = FALSE) %>%
  dplyr::arrange(Scenario_Content, Age_group, desc(avg_rating)) %>%
  dplyr::ungroup()

top3_motives_by_age 
```


## 7.3. Top 3 motives by category and age group  
```{r}
top3_motives_by_age_cat <- dat_short %>%
  dplyr::filter(Age_group %in% c("younger", "older")) %>% 
  dplyr::group_by(Scenario_Category, Age_group) %>%
  dplyr::summarise(
    across(all_of(motive_vars),
           ~ mean(.x[.x != -99], na.rm = TRUE),
           .names = "avg_{.col}"),
    .groups = "drop"
  ) %>%
  tidyr::pivot_longer(
    cols = starts_with("avg_"),
    names_to = "motive",
    values_to = "avg_rating"
  ) %>%
  dplyr::mutate(motive = sub("^avg_", "", motive)) %>%
  dplyr::group_by(Scenario_Category, Age_group) %>%
  dplyr::slice_max(avg_rating, n = 3, with_ties = FALSE) %>%
  dplyr::arrange(Scenario_Category, Age_group, desc(avg_rating)) %>%
  dplyr::ungroup()

top3_motives_by_age_cat
```

### Plot Top 3 Motives by category and age group
```{r}
ggplot(top3_motives_by_age_cat,
       aes(x = reorder(motive, avg_rating), y = avg_rating, fill = Age_group)) +
  geom_col(position = "dodge") +
  facet_wrap(~ Scenario_Category, scales = "free_x", ncol = 5) +
  coord_flip() +
  scale_fill_manual(values = c("older" = "#4CAF50", "younger" = "#2196F3"),
                    labels = c("older", "younger")) +
  labs(
    title = "Top 3 Motives by Scenario Category and Age Group",
    x = "Motive",
    y = "Mean Rating",
    fill = "Age Group"
  ) +
  theme_minimal(base_size = 12)
```


# 8.Link Decision DI to Motives 

## 8.1. Split of motives based on response DI vs. Response no DI 

### Mean Ratings per Motives -- DI 

```{r}
ratings_per_motive_DI <- dat_short %>%
  dplyr::group_by(want_to_know_bin) %>%
  dplyr::summarise(
    across(
      all_of(motive_vars),
      ~ mean(.x[.x != -99], na.rm = TRUE),
      .names = "mean_{.col}"
    ),
    .groups = "drop"
  )
```

### Highest motives and scenario -- DI
Highly rated motives where participants gave response "not want to know" (DI)
```{r}
motive_means_di <- dat_short %>%
  filter(want_to_know_bin == 1) %>%  
  group_by(Scenario_Content) %>%    
  summarise(across(
    all_of(motive_vars),
    ~ mean(.x[.x != -99], na.rm = TRUE),  
    .names = "avg_{.col}"
  )) %>%
  pivot_longer(
    cols = starts_with("avg_"),
    names_to = "motive",
    values_to = "mean_rating"
  ) %>%
  mutate(motive = sub("^avg_", "", motive)) %>%
  arrange(Scenario_Content, desc(mean_rating))

scenario_props <- dat_short %>%
  group_by(Scenario_Content) %>%
  summarise(
    prop_want_1 = mean(want_to_know_bin == 1, na.rm = TRUE),
    .groups = "drop"
  )



motive_means_di <- motive_means_di %>%
  left_join(scenario_props, by = "Scenario_Content")

motive_means_di
```

### Highest motives and Scenario -- No DI
Highly rated motives where participants gave response "want to know" 

```{r}
motive_means_noDI <- dat_short %>%
  filter(want_to_know_bin == 0) %>%  
  group_by(Scenario_Content) %>%     
  summarise(across(
    all_of(motive_vars),
    ~ mean(.x[.x != -99], na.rm = TRUE),
    .names = "avg_{.col}"
  )) %>%
  pivot_longer(
    cols = starts_with("avg_"),
    names_to = "motive",
    values_to = "mean_rating"
  ) %>%
  mutate(motive = sub("^avg_", "", motive)) %>%
  arrange(Scenario_Content, desc(mean_rating))

motive_means_noDI <- motive_means_noDI %>%
  left_join(scenario_props, by = "Scenario_Content")

motive_means_noDI
```

### Plot Motives response DI vs. no DI

```{r}
ratings_long_DI <- ratings_per_motive_DI %>%
  pivot_longer(
    cols = starts_with("mean_"),
    names_to = "motive",
    values_to = "mean_rating"
  ) %>%
  mutate(
    motive = sub("^mean_", "", motive),
    want_to_know_bin = factor(
      want_to_know_bin,
      levels = c(1, 0),
      labels = c("Not want to know (DI)", "Want to know")
    )
  )

ggplot(ratings_long_DI, aes(x = motive, y = mean_rating, group = want_to_know_bin, color = want_to_know_bin)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_color_manual(
    values = c(
      "Want to know" = "#56B4E9" ,  
      "Not want to know (DI)" =    "#E69F00"   
    )
  ) +
  theme_minimal(base_size = 12) +
  labs(
    title = "Mean Motive Ratings by DI Response",
    x = "Motive",
    y = "Mean Rating",
    color = "Response"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold", size = 14)
  )

```


## 8.2. Top 3 highest average rated motives DI vs. No DI
```{r}
top3_motives_DI <- dat_short %>%
  dplyr::group_by(want_to_know_bin) %>%
  dplyr::summarise(across(all_of(motive_vars),
                          ~ mean(.x[.x != -99], na.rm = TRUE),
                          .names = "avg_{.col}"),
                   .groups = "drop") %>%
  pivot_longer(
    cols = starts_with("avg_"),
    names_to = "motive",
    values_to = "avg_rating"
  ) %>%
  mutate(motive = sub("^avg_", "", motive)) %>%
  dplyr::group_by(want_to_know_bin) %>%
  slice_max(avg_rating, n = 3, with_ties = FALSE) %>%  
  arrange(want_to_know_bin, desc(avg_rating)) %>%
  ungroup()

top3_motives_DI
```

# 9. Link Decision DI to Motives -- young vs. old

# Top 3 motives DI vs. no DI by  across age 
```{r}
top3_motives_by_age_DI <- dat_short %>%
  dplyr::filter(Age_group %in% c("younger", "older")) %>%  
  dplyr::group_by(want_to_know_bin, Age_group) %>%
  dplyr::summarise(
    across(all_of(motive_vars),
           ~ mean(.x[.x != -99], na.rm = TRUE),
           .names = "avg_{.col}"),
    .groups = "drop"
  ) %>%
  tidyr::pivot_longer(
    cols = starts_with("avg_"),
    names_to = "motive",
    values_to = "avg_rating"
  ) %>%
  dplyr::mutate(motive = sub("^avg_", "", motive)) %>%
  dplyr::group_by(want_to_know_bin, Age_group) %>%
  dplyr::slice_max(avg_rating, n = 3, with_ties = FALSE) %>%
  dplyr::arrange(want_to_know_bin, Age_group, desc(avg_rating)) %>%
  dplyr::ungroup()

top3_motives_by_age_DI
```

### Plot Top 3 Motives by age group and DI Response
```{r}
ggplot(top3_motives_by_age_DI,
       aes(x = reorder(motive, avg_rating), y = avg_rating, fill = Age_group)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  facet_wrap(~ want_to_know_bin, scales = "free_x", ncol = 2,
             labeller = labeller(want_to_know_bin = c(`1` = "Do not want to know (DI)", `0` = "Want to know"))) +
  scale_fill_manual(values = c("older" = "#4CAF50", "younger" = "#2196F3"),
                    labels = c("older", "younger")) +
  labs(
    title = "Top 3 Motives by Age Group and DI Response",
    x = "Motive",
    y = "Average Rating",
    fill = "Age Group"
  ) +
  theme_minimal(base_size = 12)
```




