---
title: "Deliberative Ignorance"
author: "Eva Crone"
date: "2025-10-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(reshape2)
library(purrr)
library(pheatmap)
library(effsize)
library(tibble)
library(psych)
library(GPArotation)
```

# 1. Introduction

This document explores the DL2_preprocessed dataset to gain an initial understanding of the data and investigate patterns in deliberate ignorance (DI) across scenarios, motives, and age groups.

# 2. Data Preparation
## 2.1 Load Data
```{r}
dat <- read_csv("dl2_data_processed.csv")
```

## 2.2 Initial Cleaning
Exclude participants with missing DI Ratings.
```{r}
dat_short <- dat %>%
filter(!is.na(want_to_know_bin))
```

## 2.3. Include categories in dat_short
```{r}
dat_short <- dat_short %>%
  mutate(
    Scenario_Category = case_when(
      Scenario_Content %in% c(
        "hiv_test", "gen_risk_alzheimer", "gen_risk_huntington", 
        "life_expect_huntington", "spread_metast_cancer", "risk_anaesthesia",
        "nutritional_deficiencies", "calories_food", "sex_of_child"
      ) ~ "Health",
      
      Scenario_Content %in% c(
        "partner_cheating", "partner_prev_relationship",
        "converts",
        "friend_polit_opin", "friend_sex_life"
      ) ~ "Social",
      
      Scenario_Content %in% c(
        "colleague_salary", "bank_balance",
        "altern_recent_purchase", "friend_finance_problems"
      ) ~ "Finances",
      
      Scenario_Content %in% c(
        "success_chance_new_comp", "gender_job_applicant", "teaching_eval"
      ) ~ "Career",
      
      Scenario_Content %in% c(
        "hunger_crisis", "war", "holocaust_movie",
        "child_abuse", "homeless_ppl"
      ) ~ "Society",
      
      Scenario_Content %in% c(
        "Kardashian_home", "ending_GoT"
      ) ~ "Entertainment",
      
      Scenario_Content %in% c(
        "global_warming", "factory_farming"
      ) ~ "Environment",
    )
  )

```


## 2.4. Create Wide Format 
```{r}
motive_vars <- c("expectation","valence","arousal", "sadness","anger","fear", 
                  "jealousy","envy","disgust","regret","feelinggood","relief",
                  "happiness","complexity","relevance","reliability","relationships",
                  "conflict","politeness","notlying","unsureresp.","exclusion",
                  "self.percept.","actiongeneral","actionnow","disadvantage",
                  "goalinterfer","socialobligat","legalobligat","surprise","fairness")


data_wide <- dat_short %>%
  select(ResponseId, Age, Age_group, Scenario_Code, all_of(motive_vars)) %>%
  # long -> wide
  pivot_wider(
    id_cols = c(ResponseId, Age, Age_group), 
    names_from = Scenario_Code,
    values_from = all_of(motive_vars),
    names_sep = "_"
  )

```


# 3. Participants Overview 

## 3.1. Participants in total
```{r}
nrow(data_wide)
```

## 3.2. Participants count young and old

```{r}
data_wide %>%
  dplyr::group_by(Age_group) %>%
  dplyr::summarise(count = n())
```

## 3.3. Age range 
```{r}
range(data_wide$Age, na.rm = TRUE)
```
### Plot Age range
```{r}
hist(data_wide$Age,
     main = "Age range",
     xlab = "Age",
     ylab = "Number of Participants",
     col = "skyblue",
     border = "white")
```


## 3.4. Ratings per participant

### Number of DI ratings per participant
```{r}
ratings_count_person <- dat_short %>%
  dplyr::filter(want_to_know_bin %in% c(0, 1)) %>%  
  dplyr::group_by(ResponseId) %>%
  dplyr::summarise(
    n_valid_ratings = n(),   
    .groups = "drop"
  )
```

### Number of Motive Ratings per Participant
```{r}
rating_cols <- setdiff(names(data_wide), c("ResponseId", "Age_group"))

motives_count_person <- data_wide %>%
  mutate(
    n_ratings_filled = rowSums(
      select(., all_of(rating_cols)) != -99 & !is.na(select(., all_of(rating_cols)))
    )
  ) %>%
  select(ResponseId, n_ratings_filled)

### Combine Both Counts
ratings_combined <- merge(
  ratings_count_person,
  motives_count_person,
  by = "ResponseId"
)

ratings_combined <- ratings_combined %>%
  mutate(avg_ratings_per_scenario = n_ratings_filled / n_valid_ratings)
```

### Overall mean rating per scenario
```{r}
mean(ratings_combined$avg_ratings_per_scenario, na.rm = TRUE)
```


## 3.5. Number of ratings by scenario and age group
```{r}
dat_short <- dat_short %>%
  mutate(Age_group = ifelse(is.na(Age_group), "unknown", Age_group))

scenario_participants <- dat_short %>%
  dplyr::group_by(Scenario_Content, Age_group) %>%
  dplyr::summarise(n_ratings = n(), .groups = "drop")

scenario_participants
```

# 4. Deliberative Ignorance 
Responses Want to Know / Not Want to Know (DI) 

## 4.1 Overall proportion of DI across all scenarios
```{r}
overall_DI <- dat_short %>%
  dplyr::filter(!is.na(want_to_know_bin)) %>%
  dplyr::summarise(
    total_responses = n(),
    total_DI = sum(want_to_know_bin == 1),
    proportion_DI = mean(want_to_know_bin == 1)
  )

overall_DI 
```

## 4.2. DI Ratings by scenario 
### How many participants in general answered the scenario
```{r}
scenario_counts <- dat_short %>%
  dplyr::group_by(Scenario_Content) %>%
  dplyr::summarise(n_ratings = n(), .groups = "drop")

scenario_counts
```


### Count of want to know = 1 (DI) per scenario
```{r}
di_scenario <- dat_short %>% 
  dplyr::group_by(Scenario_Content, want_to_know_bin) %>% 
  dplyr::summarise(n_ratings = n(),.groups = "drop")

```

### Plot DI responses by scenario
```{r pressure, echo=FALSE}
ggplot(di_scenario, aes(x = Scenario_Content, y = n_ratings, fill = factor(want_to_know_bin))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("0" = "skyblue", "1" = "#E69F00"),
                    name = "Want to know?",
                    labels = c("0 = want to know", "1 = do NOT want to know (DI)")) +
  labs(
    title = "Want-to-Know Responses by Scenario",
    x = "Scenario",
    y = "Number of Ratings"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold", hjust = 0.5)
  )

```


## 4.3. Proportion DI vs. no DI by scenario
```{r}
di_scenario_full <- dat_short %>% 
  dplyr::group_by(Scenario_Content) %>%
  dplyr::summarise(
    n_0  = sum(want_to_know_bin == 0, na.rm = TRUE),
    n_di  = sum(want_to_know_bin == 1, na.rm = TRUE),
    total = n_0 + n_di,
    prop_1 = n_di / total,
    .groups = "drop"
  )%>%
  dplyr::arrange(desc(prop_1))

di_scenario_full <- di_scenario_full %>%
  mutate(Scenario_Content = factor(Scenario_Content, levels = Scenario_Content[order(prop_1)]))
```

### Plot proportion DI vs. no DI by scenario
```{r}
ggplot(di_scenario_full, aes(x = Scenario_Content, y = prop_1)) +
  geom_col(fill = "#E69F00") +
  coord_flip() +
  labs(
    title = "Proportion of DI per Scenario",
    x = "Scenario",
    y = "Proportion of DI to total"
  ) +
  theme_minimal(base_size = 14)
```

## 4.4. Count DI vs. no DI by category 
```{r}
di_category <- dat_short %>% 
  dplyr::group_by(Scenario_Category, want_to_know_bin) %>% 
  dplyr::summarise(n_ratings = n(),.groups = "drop")
```

### Plot DI count by category
```{r}
ggplot(di_category, aes(x = Scenario_Category, y = n_ratings, fill = factor(want_to_know_bin))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("0" = "skyblue", "1" = "#E69F00"),
                    name = "Want to know?",
                    labels = c("0 = want to know", "1 = do NOT want to know (DI)")) +
  labs(
    title = "Want-to-Know Responses by Category",
    x = "Category",
    y = "Number of Ratings"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold", hjust = 0.5)
  )

```


## 4.5.Proportion DI by category
```{r}
di_category_full <- dat_short %>% 
  dplyr::group_by(Scenario_Category) %>%
  dplyr::summarise(
    n_0  = sum(want_to_know_bin == 0, na.rm = TRUE),
    n_di  = sum(want_to_know_bin == 1, na.rm = TRUE),
    total = n_0 + n_di,
    prop_1 = n_di / total,
    .groups = "drop"
  )%>%
  dplyr::arrange(desc(prop_1))

di_category_full <- di_category_full %>%
  mutate(Scenario_Category = factor(Scenario_Category, levels = Scenario_Category[order(prop_1)]))
```

### Plot proportion DI vs. no DI by category
```{r}
ggplot(di_category_full, aes(x = Scenario_Category, y = prop_1)) +
  geom_col(fill = "#E69F00") +
  coord_flip() +
  labs(
    title = "Proportion of DI per Category",
    x = "Category",
    y = "Proportion of DI to total"
  ) +
  theme_minimal(base_size = 14)
```

# 5. Deliberative Ignorance — young vs. old 
Decisions Want to Know / Not to Know (DI)

## 5.1. Difference overall proportion -- young vs. old 
```{r}
overall_DI_age <- dat_short %>%
  dplyr::filter(!is.na(want_to_know_bin)) %>%
  dplyr::group_by(Age_group) %>%        
  dplyr::summarise(
    total_responses = n(),
    total_DI = sum(want_to_know_bin == 1),
    proportion_DI = mean(want_to_know_bin == 1),
    .groups = "drop"
  )
overall_DI_age
```

## 5.2. Differences DI by scenario --  young vs. old 
positive: more young participants chose “not want to know” 
```{r}
scenario_age_prop <- dat_short %>%
  dplyr::group_by(Scenario_Content, Age_group) %>%
  dplyr::summarise(
    total_participants = n(),
    n_want_to_know = sum(want_to_know_bin == 1, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::mutate(prop_want_to_know = n_want_to_know / total_participants) %>%
  dplyr::select(-n_want_to_know, -total_participants) %>%
  pivot_wider(
    names_from = Age_group,
    values_from = prop_want_to_know,
    values_fill = 0
  ) %>%
  mutate(diff_prop_young_old = younger - older) %>%
  arrange(desc(diff_prop_young_old)) 

scenario_age_prop 
```


### Plot Differences DI by scenario and age group

```{r}
scenario_age_long <- scenario_age_prop %>%
  pivot_longer(cols = c(younger, older),
               names_to = "Age_group",
               values_to = "prop_want_to_know") %>%
  group_by(Scenario_Content) %>%
  mutate(mean_prop = mean(prop_want_to_know, na.rm = TRUE)) %>%
  ungroup()

ggplot(scenario_age_long,
       aes(x = reorder(Scenario_Content, mean_prop),
           y = prop_want_to_know,
           fill = Age_group)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("older" = "#4CAF50", "younger" = "#2196F3"),
                    labels = c("older", "younger")) +
  labs(
    title = "DI ratings by scenario and age group",
    x = "Scenario",
    y = "Proportion DI",
    fill = "Age group"
  ) +
  coord_flip() +
  theme_minimal(base_size = 13)
```


## 5.3. Differences in DI by category -- young vs. old
positive: more young participants chose “not want to know” 
```{r}
category_age_prop <- dat_short %>%
  dplyr::group_by(Scenario_Category, Age_group) %>%
  dplyr::summarise(
    total_participants = n(),
    n_want_to_know = sum(want_to_know_bin == 1, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::mutate(prop_want_to_know = n_want_to_know / total_participants) %>%
  dplyr::select(-n_want_to_know, -total_participants) %>%
  pivot_wider(
    names_from = Age_group,
    values_from = prop_want_to_know,
    values_fill = 0
  ) %>%
  mutate(diff_prop_young_old = younger - older) %>%
  arrange(desc(diff_prop_young_old)) 

category_age_prop
```

### Plot Differences DI by category and age group

```{r}
scenario_age_long_cat <- category_age_prop %>%
  pivot_longer(cols = c(younger, older),
               names_to = "Age_group",
               values_to = "prop_want_to_know") %>%
  group_by(Scenario_Category) %>%
  mutate(mean_prop = mean(prop_want_to_know, na.rm = TRUE)) %>%
  ungroup()

ggplot(scenario_age_long_cat,
       aes(x = reorder(Scenario_Category, mean_prop),
           y = prop_want_to_know,
           fill = Age_group)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("older" = "#4CAF50", "younger" = "#2196F3"),
                    labels = c("older", "younger")) +
  labs(
    title = "DI ratings by category and age group",
    x = "Category",
    y = "Response Not want to know (DI)",
    fill = "Age group"
  ) +
  coord_flip() +
  theme_minimal(base_size = 13)
```




# 6.Ratings of Motives

## 6.1. Overall highly rated motives
```{r}
overall_motives <- dat_short %>%
  summarise(across(all_of(motive_vars), ~ mean(.x[.x != -99], na.rm = TRUE))) %>%
  pivot_longer(cols = everything(), names_to = "motive", values_to = "mean_rating") %>%
  arrange(desc(mean_rating))

overall_motives
```

## 6.2. Mean Ratings of motives by scenario
```{r}
ratings_per_motive_mean <- dat_short %>%
  dplyr::group_by(Scenario_Content) %>%
  dplyr::summarise(
    across(
      all_of(motive_vars),
      ~ mean(.x[.x != -99], na.rm = TRUE),
      .names = "mean_{.col}"
    ),
    .groups = "drop"
  )
```


### Plot: Heatmap mean ratings of motives by scenario
```{r}
ratings_long <- ratings_per_motive_mean %>%
  pivot_longer(
    cols = starts_with("mean_"),
    names_to = "Motive",
    values_to = "MeanRating"
  )


ggplot(ratings_long, aes(x = Motive, y = Scenario_Content, fill = MeanRating)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "steelblue") +
  labs(
    title = "Mean Ratings of Motives per Scenario",
    x = "Motive",
    y = "Scenario",
    fill = "Mean Rating"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

## 6.3. Top 3 highest average rated motives by scenario
```{r}
top3_motives <- dat_short %>%
  dplyr::group_by(Scenario_Content) %>%
  dplyr::summarise(across(all_of(motive_vars),
                   ~ mean(.x[.x != -99], na.rm = TRUE),  
                   .names = "avg_{.col}"),
            .groups = "drop") %>%
  pivot_longer(
    cols = starts_with("avg_"),
    names_to = "motive",
    values_to = "avg_rating"
  ) %>%
  mutate(motive = sub("^avg_", "", motive)) %>%
  dplyr::group_by(Scenario_Content) %>%
  slice_max(avg_rating, n = 3, with_ties = FALSE) %>% 
  arrange(Scenario_Content, desc(avg_rating)) %>%
  ungroup()
```


### Table: Top 3 average motives per scenario 
```{r}
top3_motives_table <- dat_short %>%
  dplyr::group_by(Scenario_Content) %>%
  dplyr::summarise(
    across(
      all_of(motive_vars),
      ~ mean(.x[.x != -99], na.rm = TRUE),  
      .names = "avg_{.col}"
    ),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = starts_with("avg_"),
    names_to = "motive",
    values_to = "avg_rating"
  ) %>%
  mutate(motive = sub("^avg_", "", motive)) %>%
  dplyr::group_by(Scenario_Content) %>%
  slice_max(avg_rating, n = 3, with_ties = FALSE) %>%
  arrange(Scenario_Content, desc(avg_rating)) %>%
  mutate(rank = row_number()) %>%
  pivot_wider(
    names_from = rank,
    values_from = c(motive, avg_rating),
    names_glue = "Top{rank}_{.value}"
  ) %>%
  arrange(Scenario_Content)

knitr::kable(top3_motives_table, digits = 2, caption = "Top 3 Motives (Mean Ratings) per Scenario")
```

### Plot Top 3 Motives by Scenario

```{r}
ggplot(top3_motives, aes(x = motive, y = Scenario_Content, fill = avg_rating)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "steelblue") +
  labs(
    title = "Top 3 Motives by Scenario",
    x = "Motive",
    y = "Scenario",
    fill = "Mean Rating"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 6.4. Mean Ratings of motives by category 

```{r}
ratings_per_motive_mean_cat <- dat_short %>%
  dplyr::group_by(Scenario_Category) %>%
  dplyr::summarise(
    across(
      all_of(motive_vars),
      ~ mean(.x[.x != -99], na.rm = TRUE),
      .names = "mean_{.col}"
    ),
    .groups = "drop"
  )

ratings_per_motive_mean_cat
```

### Plot: Heatmap mean ratings of motives by category
```{r}
ratings_long_cat <- ratings_per_motive_mean_cat %>%
  pivot_longer(
    cols = starts_with("mean_"),
    names_to = "Motive",
    values_to = "MeanRating"
  )


ggplot(ratings_long_cat, aes(x = Motive, y = Scenario_Category, fill = MeanRating)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "steelblue") +
  labs(
    title = "Mean Ratings of Motives by Category",
    x = "Motive",
    y = "Category",
    fill = "Mean Rating"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```



## 6.5. Top 3 highest average rated motives by category
```{r}
top3_motives_cat <- dat_short %>%
  dplyr::group_by(Scenario_Category) %>%
  dplyr::summarise(across(all_of(motive_vars),
                          ~ mean(.x[.x != -99], na.rm = TRUE),
                          .names = "avg_{.col}"),
                   .groups = "drop") %>%
  pivot_longer(
    cols = starts_with("avg_"),
    names_to = "motive",
    values_to = "avg_rating"
  ) %>%
  mutate(motive = sub("^avg_", "", motive)) %>%
  dplyr::group_by(Scenario_Category) %>%
  slice_max(avg_rating, n = 3, with_ties = FALSE) %>% 
  arrange(Scenario_Category, desc(avg_rating)) %>%
  ungroup()
```

### Table: Top3 motives by category
```{r}
top3_motives_cat_table <- dat_short %>%
  dplyr::group_by(Scenario_Category) %>%
  dplyr::summarise(
    across(
      all_of(motive_vars),
      ~ mean(.x[.x != -99], na.rm = TRUE),  
      .names = "avg_{.col}"
    ),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = starts_with("avg_"),
    names_to = "motive",
    values_to = "avg_rating"
  ) %>%
  mutate(motive = sub("^avg_", "", motive)) %>%
  dplyr::group_by(Scenario_Category) %>%
  slice_max(avg_rating, n = 3, with_ties = FALSE) %>%
  arrange(Scenario_Category, desc(avg_rating)) %>%
  mutate(rank = row_number()) %>%
  pivot_wider(
    names_from = rank,
    values_from = c(motive, avg_rating),
    names_glue = "Top{rank}_{.value}"
  ) %>%
  arrange(Scenario_Category)


knitr::kable(top3_motives_cat_table, digits = 2, caption = "Top 3 Motives (Mean Ratings) per Category")
```


### Plot Top 3 motives by category
```{r}
ggplot(top3_motives_cat, aes(x = motive, y = Scenario_Category, fill = avg_rating)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "steelblue") +
  labs(
    title = "Top 3 Motives by Category",
    x = "Motive",
    y = "Category",
    fill = "Mean Rating"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```



# 7. Ratings of Motives -- young vs. old 

## 7.1. Overall highly rated motives -- young vs. old 
```{r}
overall_motives_by_age <- dat_short %>%
  dplyr::filter(Age_group %in% c("younger", "older")) %>%   
  dplyr::group_by(Age_group) %>%
  dplyr::summarise(
    across(all_of(motive_vars),
           ~ mean(.x[.x != -99], na.rm = TRUE),
           .names = "avg_{.col}")
  ) %>%
  tidyr::pivot_longer(
    cols = starts_with("avg_"),
    names_to = "motive",
    values_to = "mean_rating"
  ) %>%
  dplyr::mutate(motive = sub("^avg_", "", motive)) %>%
  dplyr::arrange(Age_group, desc(mean_rating))

overall_motives_by_age
```

### Plot Overall motive ratings by age group
```{r}
ggplot(overall_motives_by_age, aes(x = reorder(motive, mean_rating), 
                                   y = mean_rating, 
                                   fill = Age_group)) +
  geom_col(position = position_dodge(width = 0.8)) +
  coord_flip() +
  scale_fill_manual(values = c("older" = "#4CAF50", "younger" = "#2196F3"),
                    labels = c("older", "younger")) +
  labs(
    title = "Overall Motive Ratings by Age Group and Scenario",
    x = "Motive",
    y = "Average Rating",
    fill = "Age Group"
  ) +
  theme_minimal(base_size = 12)
```



## 7.2. Top 3 motives by scenario - young vs. old  
```{r}
top3_motives_by_age <- dat_short %>%
  dplyr::filter(Age_group %in% c("younger", "older")) %>%  
  dplyr::group_by(Scenario_Content, Age_group) %>%
  dplyr::summarise(
    across(all_of(motive_vars),
           ~ mean(.x[.x != -99], na.rm = TRUE),
           .names = "avg_{.col}"),
    .groups = "drop"
  ) %>%
  tidyr::pivot_longer(
    cols = starts_with("avg_"),
    names_to = "motive",
    values_to = "avg_rating"
  ) %>%
  dplyr::mutate(motive = sub("^avg_", "", motive)) %>%
  dplyr::group_by(Scenario_Content, Age_group) %>%
  dplyr::slice_max(avg_rating, n = 3, with_ties = FALSE) %>%
  dplyr::arrange(Scenario_Content, Age_group, desc(avg_rating)) %>%
  dplyr::ungroup()

top3_motives_by_age 
```


## 7.3. Top 3 motives by category and age group  
```{r}
top3_motives_by_age_cat <- dat_short %>%
  dplyr::filter(Age_group %in% c("younger", "older")) %>% 
  dplyr::group_by(Scenario_Category, Age_group) %>%
  dplyr::summarise(
    across(all_of(motive_vars),
           ~ mean(.x[.x != -99], na.rm = TRUE),
           .names = "avg_{.col}"),
    .groups = "drop"
  ) %>%
  tidyr::pivot_longer(
    cols = starts_with("avg_"),
    names_to = "motive",
    values_to = "avg_rating"
  ) %>%
  dplyr::mutate(motive = sub("^avg_", "", motive)) %>%
  dplyr::group_by(Scenario_Category, Age_group) %>%
  dplyr::slice_max(avg_rating, n = 3, with_ties = FALSE) %>%
  dplyr::arrange(Scenario_Category, Age_group, desc(avg_rating)) %>%
  dplyr::ungroup()

top3_motives_by_age_cat
```

### Plot Top 3 Motives by category and age group
```{r}
ggplot(top3_motives_by_age_cat,
       aes(x = reorder(motive, avg_rating), y = avg_rating, fill = Age_group)) +
  geom_col(position = "dodge") +
  facet_wrap(~ Scenario_Category, scales = "free_x", ncol = 5) +
  coord_flip() +
  scale_fill_manual(values = c("older" = "#4CAF50", "younger" = "#2196F3"),
                    labels = c("older", "younger")) +
  labs(
    title = "Top 3 Motives by Category and Age Group",
    x = "Motive",
    y = "Mean Rating",
    fill = "Age Group"
  ) +
  theme_minimal(base_size = 12)
```


# 8.Link Decision DI to Motives 

## 8.1. Split of motives based on response DI vs. Response no DI 

### Mean Ratings per Motives -- DI 

```{r}
ratings_per_motive_DI <- dat_short %>%
  dplyr::group_by(want_to_know_bin) %>%
  dplyr::summarise(
    across(
      all_of(motive_vars),
      ~ mean(.x[.x != -99], na.rm = TRUE),
      .names = "mean_{.col}"
    ),
    .groups = "drop"
  )
```

### Highest motives and scenario -- DI
Highly rated motives where participants gave response "not want to know" (DI)
```{r}
motive_means_di <- dat_short %>%
  filter(want_to_know_bin == 1) %>%  
  group_by(Scenario_Content) %>%    
  summarise(across(
    all_of(motive_vars),
    ~ mean(.x[.x != -99], na.rm = TRUE),  
    .names = "avg_{.col}"
  )) %>%
  pivot_longer(
    cols = starts_with("avg_"),
    names_to = "motive",
    values_to = "mean_rating"
  ) %>%
  mutate(motive = sub("^avg_", "", motive)) %>%
  arrange(Scenario_Content,desc(mean_rating))

scenario_props <- dat_short %>%
  group_by(Scenario_Content) %>%
  summarise(
    prop_want_1 = mean(want_to_know_bin == 1, na.rm = TRUE),
    .groups = "drop"
  )



motive_means_di <- motive_means_di %>%
  left_join(scenario_props, by = "Scenario_Content")

motive_means_di
```

### Highest motives and Scenario -- No DI
Highly rated motives where participants gave response "want to know" 

```{r}
motive_means_noDI <- dat_short %>%
  filter(want_to_know_bin == 0) %>%  
  group_by(Scenario_Content) %>%     
  summarise(across(
    all_of(motive_vars),
    ~ mean(.x[.x != -99], na.rm = TRUE),
    .names = "avg_{.col}"
  )) %>%
  pivot_longer(
    cols = starts_with("avg_"),
    names_to = "motive",
    values_to = "mean_rating"
  ) %>%
  mutate(motive = sub("^avg_", "", motive)) %>%
  arrange(Scenario_Content, desc(mean_rating))

motive_means_noDI <- motive_means_noDI %>%
  left_join(scenario_props, by = "Scenario_Content")

motive_means_noDI
```

### Plot Motives response DI vs. no DI

```{r}
ratings_long_DI <- ratings_per_motive_DI %>%
  pivot_longer(
    cols = starts_with("mean_"),
    names_to = "motive",
    values_to = "mean_rating"
  ) %>%
  mutate(
    motive = sub("^mean_", "", motive),
    want_to_know_bin = factor(
      want_to_know_bin,
      levels = c(1, 0),
      labels = c("Not want to know (DI)", "Want to know")
    )
  )

ggplot(ratings_long_DI, aes(x = motive, y = mean_rating, group = want_to_know_bin, color = want_to_know_bin)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_color_manual(
    values = c(
      "Want to know" = "#56B4E9" ,  
      "Not want to know (DI)" =    "#E69F00"   
    )
  ) +
  theme_minimal(base_size = 12) +
  labs(
    title = "Mean Motive Ratings by DI Response",
    x = "Motive",
    y = "Mean Rating",
    color = "Response"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold", size = 14)
  ) 

```


### Bar Plot Motives response DI vs. no DI

```{r}

mean_motive_DI <- dat_short %>%
  group_by(want_to_know_bin) %>%
  summarise(
    across(all_of(motive_vars), 
           list(
             mean = ~mean(ifelse(.x == -99, NA, .x), na.rm = TRUE),
             sd   = ~sd(ifelse(.x == -99, NA, .x), na.rm = TRUE)
           ),
           .names = "{.col}_{.fn}"),
    n = n(),
    .groups = "drop"
  )

mean_motive_DI


# long format 
mean_motive_DI_long <- mean_motive_DI %>%
  pivot_longer(
    cols = -c(want_to_know_bin, n),
    names_to = c("motive", "stat"),
    names_sep = "_",
    values_to = "value"
  ) %>%
  filter(stat == "mean") 

# Barplot
ggplot(mean_motive_DI_long, aes(x = reorder(motive, value), y = value, fill = factor(want_to_know_bin))) +
  geom_col(position = position_dodge(width = 0.8)) +
  coord_flip() +
  scale_fill_manual(values = c("0" = "skyblue", "1" = "#E69F00"), name = "DI Group") +
  labs(x = "Motive", y = "Mean rating", title = "Mean Ratings per Motive DI vs. no DI") +
  theme_minimal()

```





## 8.2. Motive Ratings by Scenario and DI Response

### for schleife: plot every scenario 
```{r}

all_scenarios <- unique(dat_short$Scenario_Content)

for (sc in all_scenarios) {
  
  cat("\n### Plot for scenario:", sc, "###\n")
  
  dat_sc <- dat_short %>%
    filter(Scenario_Content == sc)
  
  # DI: not want to know (1)
  motive_means_di_all <- dat_sc %>%
    filter(want_to_know_bin == 1) %>%
    summarise(across(
      all_of(motive_vars),
      ~ mean(.x[.x != -99], na.rm = TRUE),
      .names = "mean_{.col}"
    )) %>%
    pivot_longer(starts_with("mean_"),
                 names_to = "motive",
                 values_to = "mean_rating") %>%
    mutate(
      motive = sub("^mean_", "", motive),
      want_to_know_bin = "Not want to know (DI)"
    )
  
  # want to know (0)
  motive_means_noDI_all <- dat_sc %>%
    filter(want_to_know_bin == 0) %>%
    summarise(across(
      all_of(motive_vars),
      ~ mean(.x[.x != -99], na.rm = TRUE),
      .names = "mean_{.col}"
    )) %>%
    pivot_longer(starts_with("mean_"),
                 names_to = "motive",
                 values_to = "mean_rating") %>%
    mutate(
      motive = sub("^mean_", "", motive),
      want_to_know_bin = "Want to know"
    )
  
  # combine
  ratings_long_all <- bind_rows(motive_means_di_all, motive_means_noDI_all)
  
  # plot
  print(
    ggplot(ratings_long_all, aes(motive, mean_rating, color = want_to_know_bin, group = want_to_know_bin)) +
      geom_line(linewidth = 1) +
      geom_point(size = 2) +
      scale_color_manual(values = c(
        "Want to know" = "#56B4E9",
        "Not want to know (DI)" = "#E69F00"
      )) +
      theme_minimal(base_size = 12) +
      labs(
        title = paste("Mean Motive Ratings for Scenario:", sc),
        x = "Motive",
        y = "Mean Rating",
        color = "Response"
      ) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
  )
}

```
### t test between DI vs no DI Ratings

```{r}

# Alle Szenarien & Motive definieren
all_scenarios <- unique(dat_short$Scenario_Content)
all_motives <- motive_vars

# Funktion: Wilcoxon-Test + Cliff's Delta
run_wilcox_for_motive <- function(data, scenario, motive) {
  
  dat_sc <- data %>%
    filter(Scenario_Content == scenario) %>%
    filter(.data[[motive]] != -99)
  
  # Falls eine Gruppe leer ist → überspringen
  if (length(unique(dat_sc$want_to_know_bin)) < 2) {
    return(NULL)
  }
  
  # Gruppen definieren
  group_noDI <- dat_sc[[motive]][dat_sc$want_to_know_bin == 0]
  group_DI <- dat_sc[[motive]][dat_sc$want_to_know_bin == 1]
  
  # Wilcoxon (Mann-Whitney)
  w_res <- wilcox.test(group_noDI, group_DI, exact = FALSE, correct = TRUE)
  
  # Cliff's Delta
  delta <- effsize::cliff.delta(group_noDI, group_DI)
  
  tibble(
    scenario = scenario,
    motive = motive,
    mean_noDI = mean(group_noDI),
    mean_DI = mean(group_DI),
    W = w_res$statistic,
    p = w_res$p.value,
    cliffs_delta = delta$estimate
  )
}



# Schleife über Szenarien × Motive
results_wilcox <- map_dfr(
  all_scenarios,
  function(scen) {
    map_dfr(
      all_motives,
      function(mot) {
        run_wilcox_for_motive(dat_short, scenario = scen, motive = mot)
      }
    )
  }
)

results_wilcox


results_wilcox <- results_wilcox %>%
  mutate(
    # Zahlen auf 3 Nachkommastellen runden
    across(where(is.numeric), ~ round(.x, 3)),
    # Signifikanz-Sterne
    signif = case_when(
      p < 0.001 ~ "***",
      p < 0.01  ~ "**",
      p < 0.05  ~ "*",
      TRUE      ~ ""
    )
  )
results_wilcox


```
### Plots with significance 

```{r}

for (sc in all_scenarios) {
  
  cat("\n### Plot for scenario:", sc, "###\n")
  
  # Filter Scenario
  dat_sc <- dat_short %>% filter(Scenario_Content == sc)
  
  # Compute means for DI
  motive_means_di_all <- dat_sc %>%
    filter(want_to_know_bin == 1) %>%
    summarise(across(all_of(motive_vars),
                     ~ mean(.x[.x != -99], na.rm = TRUE),
                     .names = "mean_{.col}")) %>%
    pivot_longer(starts_with("mean_"),
                 names_to = "motive",
                 values_to = "mean_rating") %>%
    mutate(
      motive = sub("^mean_", "", motive),
      want_to_know_bin = "Not want to know (DI)"
    )
  
  # Compute means for Want to know
  motive_means_noDI_all <- dat_sc %>%
    filter(want_to_know_bin == 0) %>%
    summarise(across(all_of(motive_vars),
                     ~ mean(.x[.x != -99], na.rm = TRUE),
                     .names = "mean_{.col}")) %>%
    pivot_longer(starts_with("mean_"),
                 names_to = "motive",
                 values_to = "mean_rating") %>%
    mutate(
      motive = sub("^mean_", "", motive),
      want_to_know_bin = "Want to know"
    )
  
  # Combine to long format
  ratings_long_all <- bind_rows(motive_means_di_all, motive_means_noDI_all)
  
  # --- Merge significance ----
  ratings_long_all <- ratings_long_all %>%
    left_join(
      results_wilcox %>% filter(scenario == sc),
      by = c("motive" = "motive")
    )
  
  # Sterne oberhalb der Punkte platzieren
  ratings_long_all <- ratings_long_all %>%
    mutate(star_y = mean_rating + 0.15)   # etwas über dem Punkt
  
  # --- Plot ---
  print(
    ggplot(ratings_long_all,
           aes(x = motive, y = mean_rating,
               color = want_to_know_bin, group = want_to_know_bin)) +
      geom_line(linewidth = 1) +
      geom_point(size = 2) +
      geom_text(aes(y = star_y, label = signif), color = "black", size = 4) +
      scale_color_manual(values = c(
        "Want to know" = "#56B4E9",
        "Not want to know (DI)" = "#E69F00"
      )) +
      theme_minimal(base_size = 12) +
      labs(
        title = paste("Mean Motive Ratings (with significance) for:", sc),
        x = "Motive",
        y = "Mean Rating",
        color = "Response"
      ) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
  )
}

```


## 8.3. Top 3 highest average rated motives DI vs. No DI
```{r}
top3_motives_DI <- dat_short %>%
  dplyr::group_by(want_to_know_bin) %>%
  dplyr::summarise(across(all_of(motive_vars),
                          ~ mean(.x[.x != -99], na.rm = TRUE),
                          .names = "avg_{.col}"),
                   .groups = "drop") %>%
  pivot_longer(
    cols = starts_with("avg_"),
    names_to = "motive",
    values_to = "avg_rating"
  ) %>%
  mutate(motive = sub("^avg_", "", motive)) %>%
  dplyr::group_by(want_to_know_bin) %>%
  slice_max(avg_rating, n = 3, with_ties = FALSE) %>%  
  arrange(want_to_know_bin, desc(avg_rating)) %>%
  ungroup()

top3_motives_DI

ggplot(top3_motives_DI, aes(x = motive, y = avg_rating, fill = factor(want_to_know_bin))) +
  geom_col(position = position_dodge()) +
  scale_fill_manual(values = c("0" = "#56B4E9", "1" = "#E69F00"),
                    labels = c("Want to know", "Do not want to know")) +
  theme_minimal() +
  labs(
    title = "Top 3 Motives by Decision Type",
    x = "Motive",
    y = "Mean Rating",
    fill = "Decision Type"
  ) +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 0, hjust = 1))


```

# 9. Link Decision DI to Motives -- young vs. old

# Top 3 motives DI vs. no DI by  across age 
```{r}
top3_motives_by_age_DI <- dat_short %>%
  dplyr::filter(Age_group %in% c("younger", "older")) %>%  
  dplyr::group_by(want_to_know_bin, Age_group) %>%
  dplyr::summarise(
    across(all_of(motive_vars),
           ~ mean(.x[.x != -99], na.rm = TRUE),
           .names = "avg_{.col}"),
    .groups = "drop"
  ) %>%
  tidyr::pivot_longer(
    cols = starts_with("avg_"),
    names_to = "motive",
    values_to = "avg_rating"
  ) %>%
  dplyr::mutate(motive = sub("^avg_", "", motive)) %>%
  dplyr::group_by(want_to_know_bin, Age_group) %>%
  dplyr::slice_max(avg_rating, n = 3, with_ties = FALSE) %>%
  dplyr::arrange(want_to_know_bin, Age_group, desc(avg_rating)) %>%
  dplyr::ungroup()

top3_motives_by_age_DI
```

### Plot Top 3 Motives by age group and DI Response
```{r}
ggplot(top3_motives_by_age_DI,
       aes(x = reorder(motive, avg_rating), y = avg_rating, fill = Age_group)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  facet_wrap(~ want_to_know_bin, scales = "free_x", ncol = 2,
             labeller = labeller(want_to_know_bin = c(`1` = "Do not want to know (DI)", `0` = "Want to know"))) +
  scale_fill_manual(values = c("older" = "#4CAF50", "younger" = "#2196F3"),
                    labels = c("older", "younger")) +
  labs(
    title = "Top 3 Motives by Age Group and DI Response",
    x = "Motive",
    y = "Mean Rating",
    fill = "Age Group"
  ) +
  coord_flip() +
  theme_minimal(base_size = 12)
```






# 10. Correlation between Ratings 

## 10.1. Correlation Ratings by Scenario 
```{r}
corr_scenario <- dat_short %>%
  group_by(Scenario_Content) %>%
  summarise(
    cor_matrix = list(
      cor(
        mutate(across(all_of(motive_vars), ~ ifelse(.x == -99, NA, .x))),
        use = "pairwise.complete.obs"
      )
    ),
    .groups = "drop"
  )

```


```{r}
corr_plots <- list()
strong_corr_list <- list()

# Szenarien bestimmen
all_scenarios <- unique(corr_scenario$Scenario_Content)

limits <- c(-1, 1)

# Schleife über alle Szenarien
for (sc in all_scenarios) {
  
  # --- Korrelationsmatrix extrahieren ---
  mat <- corr_scenario$cor_matrix[[which(corr_scenario$Scenario_Content == sc)]]
  
  # In Long-Format bringen
  mat_long <- reshape2::melt(mat, varnames = c("Var1", "Var2"), value.name = "r")
  
  # --- Heatmap erzeugen ---
  p <- ggplot(mat_long, aes(x = Var1, y = Var2, fill = r)) +
    geom_tile(color = "white") +
    geom_text(aes(label = round(r, 2)), size = 2) +
    scale_fill_gradient2(low = "blue", mid = "white", high = "red",
                         midpoint = 0, limits = limits) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      axis.ticks = element_blank(),
      axis.title = element_blank(),
      panel.grid = element_blank()
    ) +
    ggtitle(paste("Correlation matrix:", sc))
  
  # Plot speichern
  corr_plots[[sc]] <- p
  
  # Plot direkt ausgeben
  print(p)
  
  # --- starke Korrelationen extrahieren ---
  strong_corr <- mat_long %>%
    filter(as.numeric(factor(Var1)) < as.numeric(factor(Var2))) %>%
    filter(!is.na(r)) %>%
    filter(abs(r) > 0.5) %>%
    mutate(direction = ifelse(r > 0, "positive", "negative")) %>%
    arrange(desc(abs(r)))
  
  strong_corr_list[[sc]] <- strong_corr
}

# corr_plots[["hiv_test"]]
# strong_corr_list[["hiv_test"]]
```



## 10.2. Correlation Ratings by Category 

```{r}
# Limits für die Heatmaps
limits <- c(-1, 1)

# Listen zum Speichern
corr_category_plots <- list()
strong_corr_category <- list()

# Alle Kategorien
all_categories <- unique(dat_short$Scenario_Category)

# Schleife über alle Kategorien
for (cat in all_categories) {
  
  # --- Daten filtern ---
  df_cat <- dat_short %>%
    filter(Scenario_Category == cat) %>%
    select(all_of(motive_vars))
  
  # -99 → NA
  df_cat[df_cat == -99] <- NA
  
  # --- Korrelationsmatrix ---
  cor_mat <- cor(df_cat, use = "pairwise.complete.obs")
  
  # --- In Long-Format bringen ---
  mat_long <- melt(cor_mat, varnames = c("Var1", "Var2"), value.name = "r")
  
  # --- Heatmap erzeugen ---
  p <- ggplot(mat_long, aes(x = Var1, y = Var2, fill = r)) +
    geom_tile(color = "white") +
    geom_text(aes(label = ifelse(is.na(r), "", round(r, 2))), size = 2) +
    scale_fill_gradient2(low = "blue", mid = "white", high = "red",
                         midpoint = 0, limits = limits) +
    coord_fixed() +  # quadratische Tiles
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      axis.text.y = element_text(size = 10),
      axis.ticks = element_blank(),
      axis.title = element_blank(),
      panel.grid = element_blank()
    ) +
    ggtitle(paste("Correlation matrix:", cat))
  
  # Plot speichern
  corr_category_plots[[cat]] <- p
  
  # Plot direkt ausgeben
  print(p)
  
  # --- starke Korrelationen extrahieren ---
  strong_corr <- mat_long %>%
    filter(as.numeric(factor(Var1)) < as.numeric(factor(Var2))) %>%
    filter(!is.na(r)) %>%
    filter(abs(r) > 0.5) %>%
    mutate(direction = ifelse(r > 0, "positive", "negative")) %>%
    arrange(desc(abs(r)))
  
  strong_corr_category[[cat]] <- strong_corr
}

# corr_category_plots[["Environment"]]
# strong_corr_category[["Environment"]]


```

## 10.3. Correlation Ratings Overall

```{r}
corr_overall <- dat_short %>%
  summarise(
    cor_matrix = list(
      cor(
        mutate(across(all_of(motive_vars), ~ ifelse(.x == -99, NA, .x))),
        use = "pairwise.complete.obs"
      )
    ),
    .groups = "drop"
  )

corr_overall$cor_matrix[[1]]
```


```{r}
#correlation matrix for category

mat_corr <- corr_overall$cor_matrix[[1]]

corr_overall
#long format
mat_long_corr <- melt(mat_corr, varnames = c("Var1", "Var2"), value.name = "r")


limits <- c(-1, 1) 

ggplot(mat_long_corr, aes(x = Var1, y = Var2, fill = r)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(r, 2)), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0,
                       limits = limits) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank()
  )+
  ggtitle("Correlation matrix: Overall")
```


### strong correlation overall ratings 
```{r}
strong_corr <- mat_long_corr %>%
  filter(as.numeric(factor(Var1)) < as.numeric(factor(Var2))) %>%
  filter(abs(r) > 0.5) %>%
  mutate(direction = ifelse(r > 0, "positive", "negative")) %>%
  arrange(desc(abs(r)))

```


### Plot strong corr overall 
The plot shows positive correlations above .5. Since there are only positive correlations above .5, the absolute value of r is not displayed here.
```{r}
ggplot(strong_corr, aes(x = Var1, y = Var2, fill = r)) +
  geom_tile(color = "white") +  # Felder mit Rahmen
  scale_fill_gradient2(
    mid = "white",   # 0
    high = "red",    # positive Korrelationen
    midpoint = 0,
    limits = c(0.5, 1)
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_blank(),
    panel.grid = element_blank()
  ) +
  ggtitle("Strong correlations (r > 0.5)")

```

## 10.4. Correlation Matrix by Scenario and DI Response 

### Corr matrix DI overall 
```{r}
corr_overall_DI <- dat_short %>%
  filter(want_to_know_bin == 1) %>% 
  summarise(
    cor_matrix = list(
      cor(
        mutate(across(all_of(motive_vars), ~ ifelse(.x == -99, NA, .x))),
        use = "pairwise.complete.obs"
      )
    ),
    .groups = "drop"
  )
```


## Correlation Matrix No DI overall 
```{r}
corr_overall_noDI <- dat_short %>%
  filter(want_to_know_bin == 0) %>% 
  summarise(
    cor_matrix = list(
      cor(
        mutate(across(all_of(motive_vars), ~ ifelse(.x == -99, NA, .x))),
        use = "pairwise.complete.obs"
      )
    ),
    .groups = "drop"
  )
```

### strong corr DI 
```{r}
mat_corr_DI <- corr_overall_DI$cor_matrix[[1]]


mat_long_corr_DI <- melt(mat_corr_DI, varnames = c("Var1", "Var2"), value.name = "r")

strong_corr_DI <- mat_long_corr_DI %>%
  filter(as.numeric(factor(Var1)) < as.numeric(factor(Var2))) %>%
  filter(abs(r) > 0.5) %>%
  mutate(direction = ifelse(r > 0, "positive", "negative")) %>%
  arrange(desc(abs(r)))
```


### strong corr no DI
```{r}
mat_corr_noDI <- corr_overall_noDI$cor_matrix[[1]]


mat_long_corr_noDI <- melt(mat_corr_noDI, varnames = c("Var1", "Var2"), value.name = "r")

strong_corr_noDI <- mat_long_corr_noDI %>%
  filter(as.numeric(factor(Var1)) < as.numeric(factor(Var2))) %>%
  filter(abs(r) > 0.5) %>%
  mutate(direction = ifelse(r > 0, "positive", "negative")) %>%
  arrange(desc(abs(r)))
```


## 10.5. Correlation Matrix by Scenario and DI vs. no DI  

### DI
```{r}
corr_DI <- list()
strong_DI <- list()

for (sc in all_scenarios) {
  
  # filter data 
  df_sc <- dat_short %>%
    filter(Scenario_Content == sc,
           want_to_know_bin == 1)   # DI (Not want to know)
  
  # motives
  df_motives <- df_sc %>%
    select(all_of(motive_vars))
  
  df_motives[df_motives == -99] <- NA
  
  # correlation matrix
  cor_mat <- cor(df_motives, use = "pairwise.complete.obs")
  corr_DI[[sc]] <- cor_mat
  
  # combine for plot 
  mat_long <- melt(cor_mat, varnames = c("Var1", "Var2"), value.name = "r")
  
  # plot heatmap 
  print(
    ggplot(mat_long, aes(x = Var1, y = Var2, fill = r)) +
      geom_tile(color = "white") +
      geom_text(aes(label = round(r, 2)), size = 2) +
      scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0,
                           limits = c(-1, 1)) +
      theme_minimal() +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank()
      ) +
      ggtitle(paste("Correlation matrix:", sc, "-- DI"))
  )
  
  # strong correlations
  strong_DI[[sc]] <- mat_long %>%
    filter(as.numeric(factor(Var1)) < as.numeric(factor(Var2))) %>%
    filter(abs(r) > 0.5) %>%
    mutate(direction = ifelse(r > 0, "positive", "negative")) %>%
    arrange(desc(abs(r)))
}

# Zugriff:
# corr_DI[["hiv_test"]]
# strong_DI[["hiv_test"]]


```


### no DI 

```{r}
corr_noDI <- list()
strong_noDI <- list()

for (sc in all_scenarios) {
  
  # filter data 
  df_sc <- dat_short %>%
    filter(Scenario_Content == sc,
           want_to_know_bin == 0)   # Want to know
  
  df_motives <- df_sc %>%
    select(all_of(motive_vars))
  
  df_motives[df_motives == -99] <- NA
  
  # correlation matrix 
  cor_mat <- cor(df_motives, use = "pairwise.complete.obs")
  corr_noDI[[sc]] <- cor_mat
  
  # combine for plot
  mat_long <- melt(cor_mat, varnames = c("Var1", "Var2"), value.name = "r")
  
  # plot heatmap 
  print(
    ggplot(mat_long, aes(x = Var1, y = Var2, fill = r)) +
      geom_tile(color = "white") +
      geom_text(aes(label = round(r, 2)), size = 2) +
      scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0,
                           limits = c(-1, 1)) +
      theme_minimal() +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank()
      ) +
      ggtitle(paste("Correlation matrix:", sc, "-- no DI"))
  )
  
  # strong correlations
  strong_noDI[[sc]] <- mat_long %>%
    filter(as.numeric(factor(Var1)) < as.numeric(factor(Var2))) %>%
    filter(abs(r) > 0.5) %>%
    mutate(direction = ifelse(r > 0, "positive", "negative")) %>%
    arrange(desc(abs(r)))
}

# Zugriff:
# corr_noDI[["hiv_test"]]
# strong_noDI[["hiv_test"]]

```

## 10.6. Correlation between pos. - neg. Motive 


```{r}
motives_of_interest <- c("feelinggood", "valence")

cor_results <- data.frame(Scenario = character(),
                          Correlation = numeric(),
                          stringsAsFactors = FALSE)

for (sc in unique(dat_short$Scenario_Content)) {
  
  df_sc <- dat_short %>%
    filter(Scenario_Content == sc,
           want_to_know_bin == 1) %>%  # Nur "want to know"
    select(all_of(motives_of_interest))
  
  # -99 als NA behandeln
  df_sc[df_sc == -99] <- NA
  
  # Korrelation berechnen
  cor_value <- cor(df_sc[[1]], df_sc[[2]], use = "pairwise.complete.obs")
  
  # Ergebnis speichern
  cor_results <- rbind(cor_results, data.frame(Scenario = sc,
                                               Correlation = cor_value))
}

cor_results

```
```{r}

ggplot(cor_results, aes(x = Scenario, y = Correlation)) +
  geom_col() +
  coord_flip() + 
  theme_minimal() +
  labs(title = "Correlation Feeling Good – Valence -- DI ",
       x = "Scenario",
       y = "Correlation")

```

### Correlation between feeling good and valence for each scenario by DI group

```{r}
motives_of_interest <- c("feelinggood", "valence")

cor_results <- data.frame(
  Scenario = character(),
  Group = character(),      # neue Spalte: 0 = No DI, 1 = DI
  Correlation = numeric(),
  stringsAsFactors = FALSE
)

for (sc in unique(dat_short$Scenario_Content)) {
  for (g in c(0, 1)) {   # beide Gruppen durchlaufen
    
    df_sc <- dat_short %>%
      filter(Scenario_Content == sc,
             want_to_know_bin == g) %>%
      select(all_of(motives_of_interest))
    
    # -99 als NA behandeln
    df_sc[df_sc == -99] <- NA
    
    # Korrelation berechnen (wenn genügend Daten)
    if (nrow(df_sc) > 2) {
      cor_value <- cor(df_sc[[1]], df_sc[[2]], use = "pairwise.complete.obs")
    } else {
      cor_value <- NA
    }
    
    # Ergebnis speichern
    cor_results <- rbind(cor_results,
                         data.frame(
                           Scenario = sc,
                           Group = as.character(g),
                           Correlation = cor_value
                         ))
  }
}

```


```{r}
ggplot(cor_results, aes(x = Scenario,
                        y = Correlation,
                        fill = Group)) +
  geom_col(position = position_dodge()) +
  coord_flip() +
  scale_fill_manual(values = c("0" = "#56B4E9", "1" = "#E69F00"),
                    labels = c("0" = "Want to know",
                               "1" = "Not want to know")) +
  theme_minimal() +
  labs(
    title = "Correlation Feeling Good – Valence by DI Group",
    x = "Scenario",
    y = "Correlation",
    fill = "Group"
  )
```

# 10.7. Correlations between Motives and DI Response by Scenario 


```{r}

cor_DI_plots <- list()

all_scenarios <- unique(dat_short$Scenario_Content)

# Schleife across all scenarios 
for (sc in all_scenarios) {
  
  # filter data 
  dat_sc <- dat_short %>%
    filter(Scenario_Content == sc)
  
  # calculate correlation  
  cor_DI <- sapply(motive_vars, function(x) {
    cor(dat_sc[[x]], dat_sc$want_to_know_bin, use = "pairwise.complete.obs")
  })
  
  # tibble for ggplot
  cor_DI_df <- tibble(
    rating = names(cor_DI),
    r = cor_DI
  )
  
  # create plot
  p <- ggplot(cor_DI_df, aes(x = reorder(rating, r), y = r, fill = r)) +
    geom_col() +
    scale_fill_gradient2(low = "blue", mid = "white", high = "red",
                         midpoint = 0, limits = c(-1, 1)) +
    coord_flip() +
    theme_minimal() +
    labs(
      title = paste("Correlation: Ratings vs. want_to_know_bin -", sc),
      x = "Rating",
      y = "r"
    )
  
  # save plot 
  cor_DI_plots[[sc]] <- p
  
  print(p)
}
```

# 12. Explorative Factor Analysis
### 1. preprocess data 
```{r}
dat_motives <- dat_short %>%
  select(all_of(motive_vars)) %>%
  mutate(across(everything(), ~ na_if(.x, -99)))

```

### 2. Parallel Analysis for Factor Number
```{r}
# Korrelation prüfen
cor_mat <- cor(dat_motives, use = "pairwise.complete.obs")

# Scree-Plot
fa.parallel(dat_motives, fa = "fa", n.iter = 100, main = "Parallel Analysis for Factor Number")

```

### 3. EFA 

```{r}
efa_result <- fa(dat_motives,
                 nfactors = 5,      # Anzahl Faktoren aus Parallelanalyse/Scree
                 rotate = "oblimin", # oder "varimax" für orthogonal
                 fm = "pa")          # "ml" = Maximum Likelihood, alternative: "pa" = Principal Axis

# Ergebnis anzeigen
print(efa_result, cutoff = 0.3) # Zeigt nur Ladungen >= 0.3


efa_result$loadings #Loading
efa_result$values #Eigenwerte 
efa_result$Vaccounted
```

### Ladungen extrahieren 
```{r}
# 1. Loadings aus dem psych::fa Objekt holen
L <- efa_result$loadings

# 2. In eine "echte" numeric-Matrix umwandeln
loadings_matrix <- as.matrix(unclass(L))

# 3. DataFrame erstellen
loadings_df <- as.data.frame(loadings_matrix)

# 4. Faktoren umbenennen (Beispiel)
colnames(loadings_df) <- c("negative_emotional", "moral_social_norm", 
                           "goal-directed_action", 
                           "social_comparison", "positive_affect")


loadings_df
```




